---
title: "Update the clustered sampling app"
author: "Farshad Ebrahimi"
date: "`r lubridate::now()`"
output: html_document
params:
  database: mars14_data
  write: FALSE
---

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE)
  knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

  library(DBI)
  library(RPostgreSQL)
  library(RPostgres)
  library(odbc)
  library(RODBC)
  library(dplyr)
  library(tidyr)
  library(lubridate)
  library(knitr)
  library(stats)
  library(gridExtra)
  library(grid)
  library(gtable)
  library(ggtext)
  library(ggplot2)
  library(knitr)
  library(plotly)
  library(cluster)
  library(factoextra)
  library(sf)

  
  #create negate of %in%
    `%!in%` = Negate(`%in%`)
  
  #db
  con <- dbConnect(odbc(), dsn = params$database)
  #db for SMP polygons
  gis_apps <- paste0("MSSQL:server=PWDGISSQL;",
                          "database=GIS_APPS;",
                          "UID=gisread;",
                          "PWD=gisread;")
  gso_db <- paste0("MSSQL:server=PWDGISSQL;",
                          "database=GSODB;",
                          "UID=gisread;",
                          "PWD=gisread;")

```

The aim of this analysis is to analyze SMP population using CWL monitoring data, categorize it, and identify underrepresented SMP features in our data set. We will also use clustering to group systems with similar characteristics and develop a stratified sampling script to randomly select a representative sample of systems proportional to their cluster and system type. This will ensure a comprehensive data set covering all types of SMPs with varying design metrics such as loading ratios, drainage area, and storm size management. The output of this analysis will help the data collection team with their future site selection.

```{r data gathering, include=FALSE}
   sysbdv <- dbGetQuery(con,"SELECT * FROM external.tbl_systembdv")
   unmonitored_smp <- dbGetQuery(con," SELECT * FROM fieldwork.viw_unmonitored_postcon_on")   
    
    
    #add the design metrics to the unmonitored smp list and renaming the loading ratio coloumn
    table_all <- unmonitored_smp %>%
  left_join(sysbdv, by="system_id") %>%
  mutate(loading_ratio = sys_lrtotalda_ft2 )
    

```

**-Bioinfiltration Clustering:**

```{r  clustering Bioinfiltration type systems }

clustered_Bioinfiltration <- table_all %>%
  filter(sys_modelinputcategory == "Bioinfiltration") %>%
  select(system_id, sys_impervda_ft2, loading_ratio, sys_rawstormsizemanaged_in, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_Bioinfiltration
normalized_cluster[,2:4] <- scale(clustered_Bioinfiltration[,2:4])

#Elbow method
#number of clusters
elbow_lined_fig <- fviz_nbclust(normalized_cluster[,2:4], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:4], centers = 2)

# Store the cluster assignments back into the clustering data frame object
clustered_Bioinfiltration$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments
table(clustered_Bioinfiltration$cluster)

```

```{r elbow plot Bioinfiltration, include=TRUE, echo=FALSE}
elbow_lined_fig

```

**-Bioretention (lined) clustering. Note that loading ratio field is not applicable here:**

```{r Bioretention lined clustering}

clustered_Bioreten_lined <- table_all %>%
  filter(sys_modelinputcategory == "Bioretention (lined)") %>%
  select(system_id, sys_impervda_ft2, sys_rawstormsizemanaged_in, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_Bioreten_lined
normalized_cluster[,2:3] <- scale(clustered_Bioreten_lined[,2:3])

#Elbow method 
#number of clusters
elbow_lined_fig <- fviz_nbclust(normalized_cluster[,2:3], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:3], centers = 3)

# Store the cluster assignments back into the clustering data frame object
clustered_Bioreten_lined$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments
table(clustered_Bioreten_lined$cluster)

```

```{r elbow plot Bioretention, include=TRUE, echo=FALSE}
elbow_lined_fig

```

**-Bioretention (unlined) clustering:**

```{r Bioretention unlined clustering}

clustered_Bioreten_unlined <- table_all %>%
  filter(sys_modelinputcategory == "Bioretention (unlined)") %>%
  select(system_id, sys_impervda_ft2, sys_rawstormsizemanaged_in, loading_ratio, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_Bioreten_unlined
normalized_cluster[,2:4] <- scale(clustered_Bioreten_unlined[,2:4])

#Elbow method 
#number of clusters
elbow_lined_fig <- fviz_nbclust(normalized_cluster[,2:4], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:4], centers = 2)

# Store the cluster assignments back into the clustering data frame object
clustered_Bioreten_unlined$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments
table(clustered_Bioreten_unlined$cluster)

```

```{r elbow plot Bioretention unlined, include=TRUE, echo=FALSE}
elbow_lined_fig

```

**-Subsurface infiltration clustering:**

```{r Subsurface infiltration clustering}

clustered_subsurface <- table_all %>%
  filter(sys_modelinputcategory == "Subsurface infiltration") %>%
  select(system_id, sys_impervda_ft2, sys_rawstormsizemanaged_in, loading_ratio, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_subsurface
normalized_cluster[,2:4] <- scale(clustered_subsurface[,2:4])

#Elbow method 
#number of clusters
elbow_lined_fig <-fviz_nbclust(normalized_cluster[,2:4], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:4], centers = 2)

# Store the cluster assignments back into the clustering data frame object
clustered_subsurface$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments
table(clustered_subsurface$cluster)

```

```{r elbow plot Subsurface infiltration, include=TRUE, echo=FALSE}
elbow_lined_fig

```

**-Subsurface slow release (lined) clustering:**

```{r Subsurface slow release (lined)	 clustering}

clustered_slowrel_lined <- table_all %>%
  filter(sys_modelinputcategory == "Subsurface slow release (lined)") %>%
  select(system_id, sys_impervda_ft2, sys_rawstormsizemanaged_in, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_slowrel_lined
normalized_cluster[,2:3] <- scale(clustered_slowrel_lined[,2:3])

#Elbow method 
#number of clusters
elbow_lined_fig <- fviz_nbclust(normalized_cluster[,2:3], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:3], centers = 3)

# Store the cluster assignments back into the clustering data frame object
clustered_slowrel_lined$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments

table(clustered_slowrel_lined$cluster)
```

```{r elbow plot Subsurface slow release, include=TRUE, echo=FALSE}
elbow_lined_fig

```

**-Subsurface slow release (unlined) clustering:**

```{r Subsurface slow release (unlined)	}

clustered_slowrel_unlined <- table_all %>%
  filter(sys_modelinputcategory == "Subsurface slow release (unlined)") %>%
  select(system_id, sys_impervda_ft2, sys_rawstormsizemanaged_in, loading_ratio, sys_modelinputcategory) %>%
  na.omit() %>%
  distinct()

#normalze using mean-sd standardization
normalized_cluster <- clustered_slowrel_unlined
normalized_cluster[,2:3] <- scale(clustered_slowrel_unlined[,2:3])

#Elbow method 
#number of clusters
elbow_lined_fig <- fviz_nbclust(normalized_cluster[,2:3], kmeans, method = "wss")

#k-means 
# Cluster using kmeans with five clusters
cluster_solution <- kmeans(normalized_cluster[,2:3], centers = 2)

# Store the cluster assignments back into the clustering data frame object
clustered_slowrel_unlined$cluster <-factor(cluster_solution$cluster) 

# Look at the distribution of cluster assignments
table(clustered_slowrel_unlined$cluster)

```

```{r elbow plot Subsurface slow release unlined, include=TRUE, echo=FALSE}
elbow_lined_fig

```

```{r batc id, include=FALSE}
#binding all sub-type groups into one table
sys_clustered_all <- bind_rows(clustered_Bioinfiltration[,c("system_id","cluster","sys_modelinputcategory")], clustered_Bioreten_lined[,c("system_id","cluster","sys_modelinputcategory")], clustered_Bioreten_unlined[,c("system_id","cluster","sys_modelinputcategory")], clustered_slowrel_lined[,c("system_id","cluster","sys_modelinputcategory")], clustered_slowrel_unlined[,c("system_id","cluster","sys_modelinputcategory")], clustered_subsurface[,c("system_id","cluster","sys_modelinputcategory")])

current_batch_id <- dbGetQuery(con,"SELECT max(batch_id) FROM fieldwork.tbl_clustered_systems") %>% pull

#add the batch id to the table
sys_clustered_all["batch_id"] <- current_batch_id + 1



```

```{r address mataching and writing to db, include=FALSE}
  basin <- suppressWarnings(st_read(gis_apps, "gisad.GSWIBASIN", quiet = TRUE)) 
  
  blueroof <- suppressWarnings(st_read(gis_apps, "gisad.GSWIBLUEROOF", quiet = TRUE)) 
  
  bumpout <- suppressWarnings(st_read(gis_apps, "gisad.GSWIBUMPOUT", quiet = TRUE)) 
  
  cistern <- suppressWarnings(st_read(gis_apps, "gisad.GSWICISTERN", quiet = TRUE)) 
  
  # something wrong with drainagewell 
  
  #drainagewell <- suppressWarnings(st_read(gis_apps, "GISAD.gswiDrainageWell", quiet = TRUE))
  
  greenroof <- suppressWarnings(st_read(gis_apps, "gisad.GSWIGREENROOF", quiet = TRUE))
  
  permeablepavement <- suppressWarnings(st_read(gis_apps, "gisad.GSWIPERMEABLEPAVEMENT", quiet = TRUE))
  
  planter <- suppressWarnings(st_read(gis_apps, "gisad.GSWIPLANTER", quiet = TRUE))
  
  raingarden <- suppressWarnings(st_read(gis_apps, "gisad.GSWIRAINGARDEN", quiet = TRUE))
  
  swale <- suppressWarnings(st_read(gis_apps, "gisad.GSWISWALE", quiet = TRUE))
  
  treetrench <- suppressWarnings(st_read(gis_apps, "gisad.GSWITREETRENCH", quiet = TRUE))
  
  tree <- suppressWarnings(st_read(gis_apps, "gisad.GSWITree", quiet = TRUE))
  
  trench <- suppressWarnings(st_read(gis_apps, "gisad.GSWITRENCH", quiet = TRUE)) 
  
  wetland <- suppressWarnings(st_read(gis_apps, "gisad.GSWIWETLAND", quiet = TRUE))


#gso address layer

  gso_address<- suppressWarnings(st_read(gso_db, "GSOADMIN.Extent_Layer", quiet = TRUE))
  
#bind all smps
  # Drop columns except the SMP_ID and XY (just in case for joining) Merge all SMPs
  
  basin <- basin %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  blueroof<- blueroof %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  bumpout <- bumpout %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  cistern <- cistern %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  #drainagewell <- drainagewell %>% select(SMP_ID)
  
  greenroof<- greenroof %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  permeablepavement <- permeablepavement %>%  select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  planter<- planter %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  raingarden <- raingarden %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  swale <- swale %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  treetrench <- treetrench %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  tree <- tree %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  trench<- trench %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  wetland <- wetland %>% select(SMP_ID, X_STATEPLANE, Y_STATEPLANE)
  
  # something wrong with drainagewell 
  
  # SMP <- bind_rows(basin, blueroof, bumpout, cistern, drainagewell, greenroof, permeablepavement, planter, raingarden, swale, treetrench, trench, wetland)
  
  SMP <- bind_rows(basin,tree, blueroof, bumpout, cistern, greenroof, permeablepavement, planter, raingarden, swale, treetrench, trench, wetland) 
    
  st_crs(SMP) <- 2272
  st_crs(gso_address) <- 2272
  
  #Intersecting SMPs with address layer
  intersect_smp_address <- st_intersects(SMP, gso_address)
  
  #deciphering the intersection data structure
  
  Inters_Obj <- intersect_smp_address
  
  GSI <- SMP
  
  output <- NULL
  
  df <- NULL
  
  for(i in 1:length(Inters_Obj)) {
    
    temp <- Inters_Obj[[i]]
    
    if (length(temp) > 0) {
      
      
      Address <- gso_address[temp, "Address"] 
      District <- gso_address[temp, "District"] 
      Neighborhood <- gso_address[temp, "Neighborhood"] 
      
      SMPID <- GSI [i,"SMP_ID"]
      SMPID <- SMPID %>% st_set_geometry(NULL)

      df <- data.frame(SMPID, Address, District,Neighborhood)
    
      output <- rbind(output, df ) 
      
    }
    
  } 
  
  #trimming, name changing, switching to df frin sf
  output['system_id'] <- gsub('-\\d+$','',output$SMP_ID ) 
  output_final <- output %>%
    as.data.frame() %>% 
    select(system_id, address=Address, district=District, neighborhood=Neighborhood )
  
   #joining and dropping uids-this is a one-time step to replace the existing data in the db
  clustered_systems_table_ad <- sys_clustered_all %>%
    left_join(output_final, by="system_id") %>%
    distinct() 
  
  #keep only one address per system if multiple exists
  db_table <- clustered_systems_table_ad %>%
    group_by(system_id) %>%
    mutate(address = paste(address, collapse = ', ')) %>%
    distinct()
  
  ###write to db 
    if(params$write == TRUE ){
      
  dbWriteTable (con, SQL("fieldwork.tbl_clustered_systems"), db_table, append= TRUE, row.names = FALSE)
      
    }



```
